\paperTitle

### 2014-01-31 doc/icml2014/discussion.tex
done

% Summarize paper
For latent-variable models,
there exists a tension between
local optimization of likelihood,
which is broadly applicable but has no theoretical guarantees,
and the method of moments, which provides consistent estimators
but are limited to models with special structure.
The purpose of this work is to show that the two methods
can be used synergistically to produce consistent estimates
for a broader class of directed and undirected models.
A notable example are undirected log-linear models with hidden variables of
high treewidth.
% Classes of models

@@

% Bottleneck is key
Our approach provides consistent estimates for
models in which each hidden variable is a ``bottleneck''---that is,
it has three conditionally independent observations,
(although the dependence on the observations could be quite complex as a function of the parameters).
This bottleneck property of \citep{anandkumar13tensor}
has been exploited in many other contexts,
including Latent Dirichlet Allocation \cite{anandkumar12lda},
parsing \cite{hsu12identifiability},
noisy-or Bayesian networks \cite{halpern13noisyor},
mixture of linear regressions \cite{chaganty13regression},
and others.
We also introduce the idea of \emph{exclusive views} (\propertyref{exclusive-views}),
the key property that allows us to learn the correlation structure between hidden variables.

@@

% Identifiability
An interesting direction of study is to examine the necessity of the bottleneck
property.  Certainly, to some extent, three views are necessary to ensure
identifiability \cite{kruskal77three}, but requiring \emph{each} hidden variable to be
a bottleneck is still stronger than we would like.  However, we hope that judiciously
leveraging likelihood based methods in conjunction with the method of moments,
as we have done in this paper, can generate new hybrid techniques for estimating
an even richer class of latent-variable models.

>>>>

% Summarize paper
For latent-variable models,
there exists a tension between
local optimization of likelihood,
which is broadly applicable but has no global theoretical guarantees,
and the method of moments, which provides consistent estimators
but are limited to models with special structure.
The purpose of this work is to show that the two methods
can be used synergistically to produce consistent estimates
for a broader class of directed and undirected models.
A notable example are undirected log-linear models with hidden variables of
high treewidth.
% Classes of models

@@

% Bottleneck is key
Our approach provides consistent estimates for
models in which each hidden variable is a ``bottleneck''---that is,
it has three conditionally independent observations
(although the dependence on the observations could be quite complex as a function of the parameters).
This bottleneck property of \citet{anandkumar13tensor}
has been exploited in many other contexts,
including Latent Dirichlet Allocation \cite{anandkumar12lda},
parsing models \cite{hsu12identifiability},
noisy-or Bayesian networks \cite{halpern13noisyor},
mixture of linear regressions \cite{chaganty13regression},
and others.
We also introduce the idea of \emph{exclusive views} (\propertyref{exclusive-views}),
the key property that allows us to learn the correlation structure between hidden variables.

@@

% Identifiability
An interesting direction of study is to examine the necessity of the bottleneck
property.  Certainly, to some extent, three views are necessary to ensure
identifiability \cite{kruskal77three}, but requiring \emph{each} hidden variable to be
a bottleneck is still stronger than we would like.  We hope that by judiciously
leveraging likelihood-based methods in conjunction with the method of moments
we can generate new hybrid techniques for estimating
an even richer class of latent-variable models.

### 2014-01-31 doc/icml2014/undirected.tex
done

%However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
%  problem simplifies to,
%\begin{align}
%  \hat \theta 
%  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^\top Z_\sC - A(\theta), \label{eqn:obj}
%\end{align}
%which is concave in $\theta$.
\algorithmref{undirected} summarizes our approach.
In \sectionref{directed}, we showed that $\LearnMarginals$ is a consistent
estimator for $Z_\sC$ for any graphical model satisfying the exclusive views property (\propertyref{exclusive-views}). 
Consequently, the result is that the output of \algorithmref{undirected},
$\hat\theta$, is a consistent estimator.\footnote{Of course, the log-linear model must be identifiable
(violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical).
In general, we can identify the parameters up to kernel of $\nabla^2 A(\theta)$.}

@@

\paragraph{Pseudolikelihood.}
Unlike for directed models, in the undirected case,
graphs with high tree-width present a challenge
for even fully-supervised maximum likelihood.
This motivates looking at pseudolikelihood objective of \citep{besag75pseudo},
which optimizes the likelihood of each variable $a$ given its neighbors $\sN(a)$:
$\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})$.
In the fully-supervised setting, pseudolikelihood provides computationally efficient consistent estimates,
albeit at the cost of statistical efficiency
(coincidentally, the same high-level motivation for using method of moments in the first place).
However, if we write the objective, we will see that the log-partition function $A(\theta)$
is replaced with a sum over $A_a(\theta; x_{\sN(a)}) = \sum_a \exp(\theta^\top\phi(a))$,
which cannot be computed from $Z_\sC$, since the neighbors $\sN(a)$ in general do not form a clique.
However, we adapt our earlier algorithm to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
Of course, this method is only viable for graphs of low degree ($|\sN(a)|$ is small).
For example, in the grid model, the degree is only 4.
Note that neither sample nor computational complexity depends on the treewidth.

>>>>

%However, given the moments $\E[ \phi_\sC(\vx,\vh) ] = Z_\sC$, the
%  problem simplifies to,
%\begin{align}
%  \hat \theta 
%  &= \arg\max_{\theta} \sum_{\sC \in \sG} \theta_\sC^\top Z_\sC - A(\theta), \label{eqn:obj}
%\end{align}
%which is concave in $\theta$.
\algorithmref{undirected} summarizes our approach.
In \sectionref{directed}, we showed that $\LearnMarginals$ is a consistent
estimator for $Z_\sC$ for any graphical model satisfying the exclusive views property (\propertyref{exclusive-views}). 
Consequently, the result is that the output of \algorithmref{undirected},
$\hat\theta$, is a consistent estimator.\footnote{Of course, the log-linear model must be identifiable
(it is violated, for example, if two components of $\phi(\vx_\sC, \vh_\sC)$ were identical).
In general, we can identify the parameters up to kernel of $\nabla^2 A(\theta)$.}

@@

\paragraph{Pseudolikelihood.}
Unlike for directed models, in the undirected case,
graphs with high treewidth present a challenge
for even fully-supervised maximum likelihood due to the intractability of the log-partition function.
This motivates looking at pseudolikelihood objective of \citep{besag75pseudo} for a computationally more
efficient solution.
The pseudolikelihood objective sums over the log-probability of each variable $a$ given its neighbors $\sN(a)$:
$$L_\text{pseudo}(\theta) \eqdef \E_{x \sim \sD}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})\right].$$
In the fully-supervised setting, pseudolikelihood provides computationally efficient consistent estimates,
albeit at the cost of statistical efficiency.
Coincidentally, the same high-level motivation for using method of moments in the first place.

However, pseudolikelihood also replaces the log-partition function $A(\theta)$
with a sum over the variables: $A_a(\theta; x_{\sN(a)}) = \sum_a \exp(\theta^\top\phi(a))$,
which cannot be computed from just the hidden clique marginals $Z_\sC$,
since the neighbors $\sN(a)$ in general do not form a clique.
However, we adapt our earlier algorithm ($\LearnClique$) to estimate the marginals over all neighborhoods $Z_{\sN(a)}$.
While this method is only viable for graphs of low degree ($|\sN(a)|$ is small),
it can handle graphs with high treewidth.
For example, in the grid model, the degree is only 4.
Note that neither sample nor computational complexity depends on the treewidth.
